---
title: "STAT270 - Project 3 - Modeling with Regression"
author: "Jonathan Mason"
date: '`r Sys.Date()`'
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, warning = FALSE, message = FALSE}
# load the packages for graphing and data wrangling
library(ggplot2)
library(PASWR2)
library(car) # install the car package first, if not already installed
library(dplyr) 
library(lattice)
library(boot)
library(MASS)
```

**Note:** If you `Rmd` file submission knits you will receive total of **(5 points)**

## Case Study: Real Estate

Data and ideas for this case study come from (*Militino et al., 2004*).

**Problem 16, page 899** (**not entire problem**, only specified parts below)

The goal of this case study is to walk the user through the creation of a parsimonious multiple linear regression model that can be used to predict the total price (totalprice) of apartments by their hedonic (structural) characteristics. The data frame `VIT2005` contains several variables, and further description of the data can be found in the help file (listed below).

A data frame with 218 observations on the following 15 variables:

-   `totalprice` (the market total price (in Euros) of the apartment including garage(s) and storage room(s))

-   `area` (the total living area of the apartment in square meters)

-   `zone` (a factor indicating the neighborhood where the apartment is located with levels `Z11, Z21, Z31, Z32, Z34, Z35, Z36, Z37, Z38, Z41, Z42, Z43, Z44, Z45, Z46, Z47, Z48, Z49, Z52, Z53, Z56, Z61, and Z62`)

-   `category` (a factor indicating the condition of the apartment with levels `2A, 2B, 3A, 3B, 4A, 4B`, and `5A` ordered so that `2A` is the best and `5A` is the worst)

-   `age` (age of the apartment in years)

-   `floor` (floor on which the apartment is located)

-   `rooms` (total number of rooms including bedrooms, dining room, and kitchen)

-   `out` (a factor indicating the percent of the apartment exposed to the elements: The levels `E100, E75, E50, and E25`, correspond to complete exposure, `75`% exposure, `50`% exposure, and `25`% exposure, respectively.)

-   `conservation` (is an ordered factor indicating the state of conservation of the apartment. The levels `1A, 2A, 2B, and 3A` are ordered from best to worst conservation.)

-   `toilets` (the number of bathrooms)

-   `garage` (the number of garages)

-   `elevator` (indicates the absence (0) or presence (1) of elevators.)

-   `streetcategory` (an ordered factor from best to worst indicating the category of the street with levels `S2, S3, S4, and S5`)

-   `heating` (a factor indicating the type of heating with levels `1A, 3A, 3B, and 4A` which correspond to: no heating, low-standard private heating, high-standard private heating, and central heating, respectively.)

-   `storage` (the number of storage rooms outside of the apartment)

### Complete the parts below

**(10 pts)** **Quiz-Project 3 Pr.1**

(a) Characterize the shape, center, and spread of the variable `totalprice`.

**Solution:**

`ggplot(data = YOUR DATA, aes(x = variable to plot)) +  geom_density(fill = "your favorite color") +  theme_bw()`

YOUR CODE HERE:

```{r}
ggplot(data = VIT2005, aes(x = totalprice)) +  geom_density(fill = "green") +  theme_bw()

MD <- median(VIT2005$totalprice)
 
iqr <- IQR(VIT2005$totalprice)
 
c(MD, iqr)
```

**Observation:** **The distribution of `totalprice` is unimodal skewed right with a median of 269750 and an `IQR` of 100125.**

**(10 pts)** **Quiz-Project 3 Pr.2**

(b) Use `scatterplotMatrix()` from `car` package or `pairs()` to explore the relationships between totalprice and the numerical explanatory variables `area, age, floor, rooms, toilets, garage, elevator`, and `storage`.

`scatterplotMatrix( ~ totalprice + var1 + var2 + ... + var n, data = VIT2005)`, do not use more than 5 variables to produce input that fits the screen and can be reviewed. Use the command as many times as you need to review how `totalprice` correlates with other variables in the data.

**Solution:**

YOUR CODE HERE:

```{r}
# e.g. matrix can be produced with
scatterplotMatrix( ~ totalprice + area + age, data = VIT2005)

scatterplotMatrix( ~ totalprice + floor+ rooms, data = VIT2005)

scatterplotMatrix( ~ totalprice + toilets + garage, data = VIT2005)

scatterplotMatrix( ~ totalprice + elevator + storage, data = VIT2005)

```

Observation: **The variable `totalprice` appears to have a moderate linear relationship with `area`.\
The variable `totalprice` appears to have negative linear relationship with `age`.\
The variable `totalprice` appears to have a moderate linear relationship with `floor`.\
The variable `totalprice` appears to have a moderate linear relationship with `rooms`.\
The variable `totalprice` appears to have moderate linear relationship with `toilets`.\
The variable `totalprice` appears to have strong positive linear relationship with `garage`.\
The variable `totalprice` appears to have moderate linear relationship with `elevator`.\
The variable `totalprice` appears to have moderate linear relationship with `storage`.**

(c) **Total of (55 pts)** **Quiz-Project 3** Pr.3 to 8, 10 pts for each correctly removed variable, 5 pts to find the correlation.

Compute the correlation between `totalprice` and all of the other numerical variables. List the **three** variables in order along with their correlation coefficients that have the highest correlation with totalprice.

#### **Model (A)**: Use backward elimination to develop a model that predicts totalprice using the data frame `VIT2005`. Use a "P-value-to remove" of `5`%. Store the final model in the object `modelA`.

**(5 pts)** **Quiz-Project 3 Pr.3**

The correlation coefficients are:

```{r}
NUM <- c("area", "age", "floor", "rooms", "toilets", "garage","elevator", "storage")
COR <- cor(VIT2005[, "totalprice"], VIT2005[, NUM])
COR
```

**Observation:** **The highest three correlations with `totalprice` occur with `area` (0.8092), `toilets` (0.6876), and `rooms`(0.5256).**

**Model (A)** The functions `drop1()` and `update()` are used to create a model using backward elimination.

```{r}
# use a model where totalprice is regressed on all other variables
model.be <- lm(totalprice ~ ., data = VIT2005)

# 
drop1(model.be, test = "F")
```

**(10 pts)** (part c.1) **Quiz-Project 3 Pr.4**

Which one appears most **insignificant** (biggest `P-value`)? Drop it from the model.

E.g. If `age` is most insignificant, use the `update` function to update the model with the following code:

`model.be <- update(model.be, .~. - age)` - the first dot means we use the same response variable, the second all previously included predictors minus `age`. The again use `drop1` and so on, until all remaining variables are of significance `0.05` as specified in the directions of the problem

YOUR CODE HERE:

```{r}
model.be <- update(model.be, .~. - age)
drop1(model.be, test = "F")


```

**(10 pts)** (part c.2) **Quiz-Project 3 Pr.5**

Which one is to be dropped next?

Answer: **floor**

```{r}
model.be <- update(model.be, .~. - floor)
drop1(model.be, test = "F")

```

**(10 pts)** (part c.3) **Quiz-Project 3 Pr.6** Which one is to be dropped next?

Answer: **conservation**

```{r}

model.be <- update(model.be, .~. - conservation)
drop1(model.be, test = "F")
```

**(10 pts)** (part c.4) **Quiz-Project 3 Pr.7** Which one is to be dropped next?

Answer: **rooms**

```{r}
model.be <- update(model.be, .~. - rooms)
drop1(model.be, test = "F")
```

**(10 pts)** (part c.5) **Quiz-Project 3 Pr.8** Which one is to be dropped next?

Answer: **storage**

```{r}
model.be <- update(model.be, .~. - storage)
drop1(model.be, test = "F")
```

If all variable are significant create the object holding the optimal model as per the description of the problem using the code:

```{r}
formula(model.be)

modelA <- lm(formula(model.be), data = VIT2005)

```

**Question 1: (5 pts)** **Quiz-Project 3 Pr.9** Which variables are left in the model, list them?

Observation: **Backward elimination suggests using the variables `area`, `zone`, `category`, `out`, `toilets`, `garage`, `elevator`, `streetcorner`, and `heating` to best predict `totalprice`.**

(i) **(5 pts)** **Quiz-Project 3 Pr.10**

Compute $CV_n$, the leave-one-out cross validation error, for `modelA`. Set the seed to `5` and compute $CV_5$, the **five-fold** cross validation error, for `modelA`. The cross validation error for a generalized linear model can be computed using the `cv.glm()` from the `boot` package. Using the function `glm()` without passing a family argument is the same as using the function `lm()`. `R Code 12.3` provides a template of how to use the `cv.glm()` function. Note that $CV_n$ is returned with `cv.error$delta[1]`. To compute $CV_5$, pass the value `5` to the argument `K` inside the `cv.glm()` function.

**5 pts)** **Quiz-Project 3 Pr.11**

Compute $R^2$, $R^2_a$, the `AIC`, and the `BIC` for **Model (A)**. What is the proportion of total variability explained by **Model (A)**?

**Your Solution:**

(i) 

```{r}
modelAg <- glm(formula(model.be), data = VIT2005)

# # Fro cv.glm - the default is to set K equal to the number of observations in data which gives the usual leave-one-out cross-validation.
# 
modelAg <- glm(formula(model.be), data = VIT2005)
library(boot)
cv.error <- cv.glm(data = VIT2005, glmfit = modelAg)
cv.error$delta[1]

set.seed(5)
cv.error5 <- cv.glm(data = VIT2005, glmfit = modelAg, K=5)
cv.error5$delta[1]
```

```{r}
modelAg <- glm(formula(model.be), data = VIT2005)



set.seed(5) # use for replication purposes
cv.error5 <- cv.glm(data = VIT2005, glmfit = modelAg, K = 5)
CV5a <- cv.error5$delta[1]
CV5a

```

Observation: **The** $CV_n = 647026246$ **for Model (A), and** $CV_5 =652546038$ **for Model (A).**

(ii) Since this problem and a few more will request many goodness of fit statistics, a function called `mgof()` is written to compute the requested values. **Use it as shown below**.

```{r}

mgof <- function(model = model, data = DF, ...){
  R2a <- summary(model)$adj.r.squared
  R2 <- summary(model)$r.squared
  aic <- AIC(model)
  bic <- AIC(model, k = log(nrow(data)))
  se <- summary(model)$sigma
  form <- formula(model)
  ANS <- c(R2 = R2, R2.adj = R2a, AIC = aic, BIC = bic, SE = se)
  ANS
}

MGOF <- mgof(model = modelA, data = VIT2005)
MGOF
```

Observation: **The total proportion of variability (**$R^2$**) explained by modelA is `0.9138`.**

(d) **(10 pts)** **Quiz-Project 3 Pr.12**

Explore the residuals of the Models (A) using the function `residualPlot()` or `residualPlots()` from the package `car`. Comment on the results. (**Diagnostics: Checking the model assumptions**)

```{r}
residualPlot(modelA, main = "Model A")
```

**(d) Question 2:** Is there curvature of the residuals on the above plot?

**Observation: The residuals versus the fitted values for Model (A) have a definite curvature indicating the model is not quite adequate.**

**Extra Credit** (10 pts) **Quiz-Project 3 Pr.13**

**(e)** Use the function `boxCox()` from car to find a suitable transformation for totalprice. Build Model (E) Use backward elimination to develop a model that predicts `log(totalprice)` using the data frame `VIT2005`. Use a "P-value-to remove" of 5%. Store the final model in the object `modelE`.

For details on `boxCox()` see page 856 in text

```{r}
boxCox(modelA, lambda = seq(-0.5, 0.5, length = 200))
```

#### Observation:

**A `log` transformation is suggested for the response `totalprice` in Model (A).**

**Extra Credit** (20 pts) **Quiz-Project 3 Pr.14** - 4 pts for excluding the correct variable at each step.

#### **Model (E)**:

Use backward elimination to develop a model that predicts `log(totalprice)` using the data frame `VIT2005`. Use a "P-value-to remove" of `5`%. Store the final model in the object `modelE`.

**Model (E)** The functions `drop1()` and `update()` are used to create a model using backward elimination. (as shown at the beginning)

```{r}
# transform the response variable using natural log transformation

VITNEW <- VIT2005 %>% mutate(logtotalprice = log(totalprice))

# build the fitted full model
model.be <- lm(logtotalprice ~ ., data = VITNEW[ ,-1]) # exclude the totalprice variable since it is no longer response variable
drop1(model.be, test = "F")

```

and so on.

```{r}
model.be <- update(model.be, .~. - age)
drop1(model.be, test = "F")
```

```{r}
model.be <- update(model.be, .~. - conservation)
drop1(model.be, test = "F")
```

```{r}
model.be <- update(model.be, .~. - floor)
drop1(model.be, test = "F")
```

```{r}
model.be <- update(model.be, .~. - rooms)
drop1(model.be, test = "F")
```

```{r}
model.be <- update(model.be, .~. - streetcategory)
drop1(model.be, test = "F")
```

```{r}
formula(model.be)

modelE <- lm(formula(model.be), data = VITNEW[ ,-1])
```